### 2015-03-27 tex/abstract.tex
Brought in abstract, intro, and related work from CAV paper.

A fundamental challenge of using symbolic execution for software
analysis has been the treatment of dynamically allocated
data. Generalized symbolic execution (GSE) techniques have addressed
this challenge by materializing a concrete heap lazily during field
accesses.  These techniques however, exacerbate the path explosion
problem.  In this work, we present a summary heap approach which takes
inspiration from the static analysis domain and defines a bipartite
graph to encode multiple heaps in single representation, which, unlike
state-of-the-art static analysis approaches, (i) supports aliasing,
(ii) does not require rewrite of constraints on heap updates, and
(iii) can initialize the heap-on-the-fly. We also define a summary
heap algorithm to perform heap updates during symbolic execution on
this representation in order to mitigate the path explosion problem in
GSE techniques. Our approach is sound and complete with respect to the
properties provable by GSE. A preliminary evaluation of our
implementation in Java Pathfinder framework shows the summary heap
enables the initialization of larger, more complex heaps than
previously possible.

\begin{comment}
A fundamental challenge of using symbolic execution for software
analysis has been the treatment of dynamically allocated data.
State-of-the-art techniques have addressed this challenge through
either (a) the use of summaries that over-approximate possible heaps, or
(b) by materializing a concrete heap lazily during generalized
symbolic execution. In this work, we present a novel heap
initialization and analysis technique which takes inspiration from
both approaches and constructs precise heap summaries lazily during
symbolic execution.  Our approach is 1) \emph{scalable}: it reduces the
points of non-determinism compared to generalized symbolic execution
and explores each control-flow path only once for any given set of
isomorphic heaps, 2) \emph{precise}: at any given point during symbolic
execution, the symbolic heap represents the exact set of feasible
concrete heap structures for the program under analysis, and 3)
\emph{expressive}: the symbolic heap can represent recursive data
structures. We demonstrate the precision and scalability of our
approach by implementing it as an extension to the Symbolic PathFinder
framework for analyzing Java programs.
\end{comment}

%and by using 
%materializing objects on the concrete heap ``as needed'' and
%using non-deterministic choice points to explore each feasible
%concrete heap configuration.

%Because analysis of the materialized
%heap locations relies on concrete program semantics, the lazy
%initialization approach exacerbates the state space explosion
%problem that limits the scalability of symbolic execution.
%In this work we present a novel approach for lazy symbolic execution
%of heap manipulating software which utilizes 
%a fully symbolic heap constructed on-the-fly during symbolic execution.
%Our approach is 1) \emph{scalable} -- it does not
%create the additional points of non-determinism
%introduced by existing lazy initialization techniques and it explores
%each execution path only once for any given set of isomorphic heaps,
%2) \emph{precise} -- at any given point during symbolic execution, 
%the symbolic heap represents the exact set of feasible
%concrete heap structures for the program under analysis, and
%3) \emph{expressive} -- the symbolic heap can represent recursive data structures 
%and heaps resulting from loops and recursive control structures in the code. 
%We report on a case-study of an implementation of our technique in the
%Symbolic PathFinder tool to illustrate its scalability, precision
%and expressiveness. We also discuss how test case generation -- a common
 %use for symbolic execution results -- can benefit from symbolic
 %execution which uses a fully symbolic heap.

>>>>

A fundamental challenge of using symbolic execution for software
analysis is the treatment of dynamically allocated data. This paper
introduces a compact symbolic heap representation based on a field
sensitive points-to graph with edges annotated with disjunctive
formulas describing aliasing relationships, along with a set of
rewrite rules that describe the initialization, dereferencing, and
update to fields in the symbolic heaps. The symbolic heap supports
arbitrary recursive data structures and captures all possible heaps
that follow a common control flow path up to the point of
execution. An evaluation of a proof-of-concept implementation in the
Java Pathfinder framework is presented to demonstrate the
computational feasibility of the approach over several classical
symbolic execution benchmarks.

%A fundamental challenge of using symbolic execution for software
%analysis is the treatment of dynamically allocated data. Existing
%static analysis techniques over-approximate the potential heaps,
%dynamic symbolic execution techniques often under-approximate the
%heaps, while generalized symbolic execution leads to excessive
%case-splitting. This paper introduces a compact heap representation,
%along with heap update rules that (i) allow reads and writes of
%symbolic references, (ii) do not require constraint rewriting, and
%(iii) support arbitrary recursive data structures. As such, the
%representation and update rules capture all possible heaps that
%follow a common execution path. More importantly, in contrast to
%existing state-of-the-art techniques, the definition and update rules
%are unconstrained and not subject to bounds, such as the number of
%heap locations, aliasing, or memory partitions. An evaluation of a
%proof-of-concept implementation in the Java Pathfinder framework is
%presented to demonstrate the computational feasibility of the
%approach over several classical symbolic execution benchmarks.

### 2015-03-27 tex/definitions.tex
Brought in abstract, intro, and related work from CAV paper.

\section{Definitions}

\begin{definition}
The set of \textbf{references} $\mathcal{R}$ is defined as the set of natural numbers
 $$\mathcal{R} = \mathbb{N}$$
\end{definition}

@@

\begin{definition}
A \textbf{state transition relation} $\rightarrow_{\Phi}$ is a binary
relation $\rightarrow_{\Phi}\ \subseteq\ \mathcal{S} \times
\mathcal{S} $, which relates machine states with successor states. Two
important state transition relations are \textbf{GSE} $\rgse$ and
\textbf{symbolic} $\rsym$. Each of these use a separate relation for
initialization: $\rinit$ for GSE and $\rsum$ for symbolic. All of
these transition relations are defined in \figref{fig:lazy},
\figref{fig:lazyInit}, \figref{fig:sym}, and \figref{fig:symInit}.
\end{definition}

@@

\begin{lemma}[Successors of well-formed states]
\label{lem:succ}
Any well-formed state has at least one successor unless the expression is $\cfgnt{v}$ and the continuation is $\cfgt{end}$.
\end{lemma}
\begin{proof}
Please write me!
\end{proof}

@@

Note that in the literature it is customary to define bisimulation in
terms of a single labeled transition system, whereas for the purposes
of this paper the definition of bisimulation refers to a pair of
transition relations $\rightarrow_x$ and $\rightarrow_y$ defined by
reduction rules. Since it is possible to create a union of the two
rule systems $\rightarrow_x \cup \rightarrow_y$, and since none of the
transitions in the reduction rules in this paper are labeled, this
definition is sufficient for all of the customary properties of
bisimulation to apply. 
%For a more detailed treatment on the
%application of bisimulation to reduction rule systems see
%\cite{GSE:barbedbisimulation}.

@@

If you have a lattice $L$, and you have some 
function $F$ which transforms one point on the lattice into another, then
$F$ is an endofunction on $L$. For any endofunction $F$, a 
a post-fixed point of $F$ is any point $x$ where $x < F(x)$, and a pre-fixed 
point is any point $x$ where $F(x) < x$. The inductively defined set $F_\mathit{ind}$
is equal to the meet of all the pre-fixed points, and the conductively 
defined set $F_\mathit{coind}$ is equal to the join of all post-fixed points. 
$x \leq F_\mathit{coins}$. If $F$ is monotone, then the greatest fixed point $\mathit{gfp}(F)$ 
is equal to the coinductive set, and the least fixed point $\mathit{lfp}(F)$ is equal to the 
inductive set.

>>>>

\section{Definitions}
\label{sec:defs}
\begin{definition}
The set of \textbf{references} $\mathcal{R}$ is defined as the set of natural numbers
 $$\mathcal{R} = \mathbb{N}$$
\end{definition}

@@

\begin{definition}
A \textbf{state transition relation} $\rightarrow_{\Phi}$ is a binary
relation $\rightarrow_{\Phi}\ \subseteq\ \mathcal{S} \times
\mathcal{S} $, which relates machine states with successor states. Two
important state transition relations are \textbf{GSE} $\rgse$ and
\textbf{symbolic} $\rsym$. Each of these use a separate relation for
initialization: $\rinit$ for GSE and $\rsum$ for symbolic. All of
these transition relations are defined in \figref{fig:lazy},
\figref{fig:lazyInit}, \figref{fig:javalite-common}, \figref{fig:eqs}, \figref{fig:fHeap} and \figref{fig:symInit}.
\end{definition}

@@

\begin{lemma}[Successors of well-formed states]
\label{lem:succ}
Any well-formed state has at least one successor unless the expression is $\cfgnt{v}$ and the continuation is $\cfgt{end}$.
\end{lemma}
\begin{proof}
Proof of this lemma can be obtained by simple inspection of the rules.
\end{proof}

@@

Note that in the literature it is customary to define bisimulation in
terms of a single labeled transition system, whereas for the purposes
of this paper the definition of bisimulation refers to a pair of
transition relations $\rightarrow_x$ and $\rightarrow_y$ defined by
reduction rules. Since it is possible to create a union of the two
rule systems $\rightarrow_x \cup \rightarrow_y$, and since none of the
transitions in the reduction rules in this paper are labeled, this
definition is sufficient for all of the customary properties of
bisimulation to apply. 
For a more detailed treatment on the
application of bisimulation to reduction rule systems see
\cite{Sangiorgi:2011}.

@@

Given a lattice $L$, for any endofunction $F$ over points in $L$, 
a post-fixed point of $F$ is any point $x$ where $x < F(x)$, and a pre-fixed 
point is any point $x$ where $F(x) < x$. The inductively defined set $F_\mathit{ind}$
is equal to the meet of all the pre-fixed points, and the coinductively 
defined set $F_\mathit{coind}$ is equal to the join of all post-fixed points. 
If $F$ is monotone, then the greatest fixed point $\mathit{gfp}(F)$ 
is equal to the coinductive set, and the least fixed point $\mathit{lfp}(F)$ is equal to the 
inductive set.

### 2015-03-27 tex/heap.tex
Added a sentence about the immutability of references.
Changed bullets on list to be dashes as in the introduction.
Brought in abstract, intro, and related work from CAV paper.

The bipartite graph itself consists of references, $\cfgnt{r}$, and
constraint-location pairs, $\lp\phi\ \cfgnt{l}\rp$. The graph is
expressed in the location map, $\cfgnt{L}$, and the reference map,
$\cfgnt{R}$. As done with the environment, $\cfgnt{L}$ and $\cfgnt{R}$
are treated as partial functions where $\cfgnt{L}(r) =
\{(\phi\ \cfgnt{l})\ ...\}$ is the set of constraint-location pairs in
the heap associated with the given reference, and
$\cfgnt{R}(\cfgnt{l},\cfgnt{f}) = \cfgnt{r}$ is the reference
associated with the given location-field pair in the heap.

There are several properties of the summary heap enforced by the
rewrite rules that are invariant and important to the correctness
proofs in \secref{sec:bisim}: \emph{reference partitions},
\emph{immutability}, \emph{determinism}, \emph{type
  consistency}, and \emph{null and uninitialized}.

\emph{Reference partitions}: new references drawn from
  three distinct partitions (i) $\mathrm{init}_\cfgnt{r}()$ for references for the \emph{input heap}, (ii) $\mathrm{fresh}_\cfgnt{r}()$ for references for \emph{auxiliary literals}, and
  (iii) $\mathrm{stack}_\cfgnt{r}()$ for references for \emph{stack literals} in the expressions and
  continuations. The references strictly increase in value on each
  call, and modular arithmetic is used to determine the partition to
  which a reference belongs.
\begin{compactdesc}
\item[Input heap:] these references have two
  interpretations: one as a literal and another as a variable. As a literal it is a node in
  the summary heap. As a variable it is a term in any number of constraints that label
  edges in the summary heap. Only input heap references and the
  special reference $\cfgnt{r}_\mathit{null}$ appear in constraints
\item[Auxiliary literals:] these references are only used to define additional structure in
  the bipartite graph and are not part of any constraint and do not appear
  in any environment, expression, or continuation.
\item[Stack references:] these references are also used to define additional structure in the bipartite graph but may also appear in the environment, expression, or continuation. These are not part of any constraint.
\end{compactdesc}
Reference partitions are a critical piece to the existence proof of a
bisimilation between the summary heap algorithm and generalized
symbolic execution (see \secref{sec:bisim}).

@@

For example, if the solver assigns input heap variables $\cfgnt{r}$ and $\cfgnt{r}^\prime$
such that $\cfgnt{r} = \cfgnt{r}^\prime$ is true, then there exists
some location $\cfgnt{l}$ and constraints $\phi$ and $\phi^\prime$ such
that
\begin{compactitem}
\item $\lp\phi\ \cfgnt{l}\rp \in \cfgnt{L}(\cfgnt{r})$; and
\item $\lp\phi^\prime\ \cfgnt{l}\rp \in \cfgnt{L}(\cfgnt{r}^\prime)$; and
\item $\phi$ and $\phi^\prime$ hold in the aliasing assignments from the solver.
\end{compactitem}
In this way, the graph groups heaps with the conditions under which
those heaps exist. This representation separates the constraint
problem from the morphology problem: it is possible to update the heap
without having to rewrite constraints. Without this separation, it
would not be possible to lazily initialize heap locations along a
program path.

>>>>

The bipartite graph itself consists of references, $\cfgnt{r}$, and
constraint-location pairs, $\lp\phi\ \cfgnt{l}\rp$. Locations may be 
altered to perform updates, whereas references are immutable
once they are added to the graph. The graph is
expressed in the location map, $\cfgnt{L}$, and the reference map,
$\cfgnt{R}$. As done with the environment, $\cfgnt{L}$ and $\cfgnt{R}$
are treated as partial functions where $\cfgnt{L}(r) =
\{(\phi\ \cfgnt{l})\ ...\}$ is the set of constraint-location pairs in
the heap associated with the given reference, and
$\cfgnt{R}(\cfgnt{l},\cfgnt{f}) = \cfgnt{r}$ is the reference
associated with the given location-field pair in the heap.

There are several properties of the summary heap enforced by the
rewrite rules that are invariant and important to the correctness
proofs in \secref{sec:proofs}: \emph{reference partitions},
\emph{immutability}, \emph{determinism}, \emph{type
  consistency}, and \emph{null and uninitialized}.

\emph{Reference partitions}: new references drawn from
  three distinct partitions (i) $\mathrm{init}_\cfgnt{r}()$ for references for the \emph{input heap}, (ii) $\mathrm{fresh}_\cfgnt{r}()$ for references for \emph{auxiliary literals}, and
  (iii) $\mathrm{stack}_\cfgnt{r}()$ for references for \emph{stack literals} in the expressions and
  continuations. The references strictly increase in value on each
  call, and modular arithmetic is used to determine the partition to
  which a reference belongs.
\begin{compactdesc}
\item[Input heap:] these references have two
  interpretations: one as a literal and another as a variable. As a literal it is a node in
  the summary heap. As a variable it is a term in any number of constraints that label
  edges in the summary heap. Only input heap references and the
  special reference $\cfgnt{r}_\mathit{null}$ appear in constraints
\item[Auxiliary literals:] these references are only used to define additional structure in
  the bipartite graph and are not part of any constraint and do not appear
  in any environment, expression, or continuation.
\item[Stack references:] these references are also used to define additional structure in the bipartite graph but may also appear in the environment, expression, or continuation. These are not part of any constraint.
\end{compactdesc}
Reference partitions are a critical piece to the existence proof of a
bisimilation between the summary heap algorithm and generalized
symbolic execution (see \secref{sec:defs}).

@@

For example, if the solver assigns input heap variables $\cfgnt{r}$ and $\cfgnt{r}^\prime$
such that $\cfgnt{r} = \cfgnt{r}^\prime$ is true, then there exists
some location $\cfgnt{l}$ and constraints $\phi$ and $\phi^\prime$ such
that
\begin{compactitem}
\item[-] $\lp\phi\ \cfgnt{l}\rp \in \cfgnt{L}(\cfgnt{r})$; and
\item[-] $\lp\phi^\prime\ \cfgnt{l}\rp \in \cfgnt{L}(\cfgnt{r}^\prime)$; and
\item[-] $\phi$ and $\phi^\prime$ hold in the aliasing assignments from the solver.
\end{compactitem}
In this way, the graph groups heaps with the conditions under which
those heaps exist. This representation separates the constraint
problem from the morphology problem: it is possible to update the heap
without having to rewrite constraints. Without this separation, it
would not be possible to lazily initialize heap locations along a
program path.

### 2015-03-27 tex/intro.tex
Made lazier Lazier everywhere
Brought in abstract, intro, and related work from CAV paper.

In recent years symbolic execution has provided the basis for various
software testing and analysis techniques. Symbolic execution
systematically explores the program execution space using symbolic
input values, and for each explored path, computes constraints on the
symbolic inputs to create a \emph{ path condition}.  The path
conditions computed by symbolic execution characterize the observed
program execution behaviors and have been used as an enabling
technology for various applications, e.g., regression
analysis~\cite{backes:2012,Godefroid:SAS11,Person:FSE08,person:pldi2011,Ramos:2011,Yang:ISSTA12},
data structure repair~\cite{KhurshidETAL05RepairingStructurally},
dynamic discovery of
invariants~\cite{CsallnerETAL08DySy,Zhang:ISSTA14}, and
debugging~\cite{Ma:2011}.

Initial work on symbolic execution largely focused on checking
properties of programs with primitive
types~\cite{clarke76TSE,King:76}.  With the advent of object-oriented
languages, recent work has generalized the core ideas of symbolic
execution to enable analysis of programs containing complex data
structures with unbounded domains, i.e., data stored on the
heap~\cite{Kiasan06,Kiasan07,GSE03}.  These \emph{Generalized Symbolic
  Execution} (GSE) techniques construct the heap in a lazy manner,
deferring materialization of objects on the concrete heap until they
are needed for the analysis to proceed. The materialization creates
additional non-deterministic choice points in the symbolic execution
tree by representing the feasible heap configurations as (i) null,
(ii) an instance of a new reference of a compatible class, and (iii)
an alias to a previously initialized symbolic reference.  Symbolic
execution then follows concrete program semantics for materialized
heap locations. Although this approach enables the analysis of heap
manipulating programs, a large number of feasible concrete heap
configurations are created leading to a path explosion.

The goal of this work is to mitigate the path explosion problem in GSE
by grouping multiple heaps together and only partitioning the heaps at
points of divergence in the control flow graph. Our inspiration is
found in the domain of static analysis that uses sets of constraints
over heap locations to encode multiple heaps in a single
representation. These sets, sometimes known as \emph{value sets} or
\emph{value summaries}, allow multiple heaps to be represented
simultaneously with a higher degree of precision than afforded by
traditional techniques for shape analysis. Some of these previous
attempts, however, are unable to handle aliasing in heaps due to a
recursive definition of objects~\cite{Xie:2005}, while others require
a set of heaps as input and are unable to initialize heaps
on-the-fly~\cite{Dillig:2011,Tillmann:2008}.  As is typical in static
analysis, over approximation of the heap representations alleviates
some of these limitations but often leads to infeasible heaps. Also,
most static analysis techniques for heap updates require a rewrite of
the constraint system making it prohibitive for use in the context of
symbolic execution.

In this work, to effectively represent multiple heaps simultaneously
in the context of symbolic execution, we define a novel heap
representation and an algorithm to initialize and update the heap
on-the-fly. In essence, our summary heap approach supports aliasing,
does not require constraint rewriting on heap updates, and reduces
non-determinism in the search space during symbolic execution.

We represent the heap as a bipartite graph consisting of references
and locations. Each reference is able to point to multiple locations,
where each edge is predicated on constraints over aliasing between
references. Each location points to a single reference for a given
field. The use of a bipartite graph affords two key advantages: (i) it
allows for a non-recursive definition of objects which enables us to
support aliasing, and (ii) it does not introduce auxiliary variables
or require rewriting of non-local constraints during updates to the
heap.

The summary heap algorithm defines how the bipartite graph is updated
during lazy initialization. Unlike GSE, however, the summary heap
algorithm introduces non-determinism in the search only at points of
divergence in the control flow graph. These points of non-determinism
are at field accesses that lead to null-pointer exceptions and at
comparisons of references. The former represents a divergence due to
exceptional control flow while the latter is due to program structure.
The combination of the heap representation and the summary heap
algorithm enables us to improve the efficiency of symbolic execution
of heap manipulating programs over state-of-the-art techniques.

% The precise heap representation based on a bipartiate graph coupled
%with this algorithm to perform updates to the heap enable us to
%successfully mitigate the state space explosion problem in GSE.

\begin{comment}
In
practical terms, any property that can be proved using GSE can also be
proved using summary heaps, and, any property proved by
summary heaps can also be proved by GSE. Of course, both require a
bound on heap reference chains and loops.
\end{comment}

%Aside from the novelty of the heap summary and its application to
%symbolic execution, the new data structure appears to ameliorate
%state explosion in symbolic execution as hoped.

The summary heap algorithm is sound and complete with respect to
properties that are provable using GSE. This proof is accomplished by
showing the existence of a bisimulation between states in GSE and
states in the summary heap algorithm.  A preliminary implementation in
Java Pathfinder (JPF) shows that in general, the heap summary
algorithm improves over other state-of-the-art techniques, and in some
instances, the improvement is remarkable: two-orders of magnitude
reduction in running time. More importantly though, the heap summary
enables the initialization of larger more complex heaps than
previously possible as shown in our results.

%, described in this paper, compares the heap summary algorithm on a
%handful of benchmarks, each representing different degrees of
%initialization and branching, to GSE with lazy initialization in
%Symbolic Pathfinder (SPF) and implementation lazier\# also in
%SPF. The results

\begin{comment}
We implement the summary heap algorithm as an extension to \emph{Java
  PathFinder} (JPF) and compare it to GSE with lazy initialization in
\emph{Symbolic Pathfinder (SPF)} and an implementation of lazier\#
initialization in SPF.  The comparison shows on some examples up to a two-order of magnitude
reduction in the total time taken to explore the same state space
defined by the bound on the longest heap reference chain.  For these
examples, we show that where other GSE approaches are unable to
complete exploration within the provided time bound, the summary heap
finishes exploration in a few seconds.
\end{comment}

%shows that it can improve the scalability of symbolic execution of
%heap manipulating programs X fold over GSE-based analyses while
%maintaining precision and efficiency.

%preserves the heap
%configurations and program behaviors

%complete --> anything prove with GSE you can with heap sound -->
%preserves the same behaviors in the GSE machine.

%Our analysis improves on existing GSE approaches by reducing
%the size of the symbolic execution state space explored, thereby
%improving scalability of symboic execution for heap manipulating
%software. It improves on existing heap summarization techniques by
%enabling analysis of arbitrary recursive data structures (no initial
%heap configuration is required) while avoiding the imprecision
%introduced by these techniques.

%\nsr{a paragraph that provides a breif summary of the results.. would
%  be nice if we have numbers for lazier and lazier Sharp as well.}

@@

%\item\textbf{-} A summary heap that represents sets of concrete heaps
% without approximation or morphological restrictions.

\item\textbf{-} A non-approximate, morphologically arbitrary, and efficient 
summary heap representation.

\item\textbf{-} An algorithm to dynamically initialize, summarize, and
  update the summary heap during symbolic execution.

\item\textbf{-} A bisimulation proof establishing the soundness and 
completeness of the heap summary approach with respect to
properties provable by GSE.

\item\textbf{-} A proof-of-concept implementation and empirical study 
demonstrating the scalability of the summary heap approach
compared to other GSE approaches.

\end{compactdesc}

%\begin{compactdesc}
%
%\item\textbf{-} A summary heap as a bipartite graph that supports
%  aliasing and does not require rewriting of constraints.
%
%% to handle alias and efficient updates of constraints.
%
%\item\textbf{-} A summary heap algorithm to initialize, summarize, and
%  update the bipartiate graphs during symbolic execution.
%
%\item\textbf{-} A proof of the existence of a bisimulation relation to establish
%  that the heap summary approach is sound and complete with respect to
%  properties provable by GSE.
%
%\item\textbf{-} A proof of concept implementation and small empirical study to demonstrate the scability of the heap summary approach
%  compared to other GSE approaches.
%
%\end{compactdesc}

%The rest of this paper defines summary heap and its to symbolic
%execution. Following the algorithm definition, the existence proof of
%bisimulation is given with the accompanying corollaries for soundness
%and completeness. This is followed by the implementation in JPF and
%the evaluation. The paper ends with the conclusion and future work.

%\item{We present a novel inter-procedural analysis technique for 
%symbolic execution of heap manipulating software which constructs 
%a precise heap summary.}
%\item{We use the JavaLite language~\cite{saints-MS} to formalize 
%the algorithm for constructing and manipulating the heap summary
%on-the-fly during symbolic execution.}
%\item{We implement our approach in the Symbolic PathFinder tool and
%empirically demonstrate that our technique improves
%the scalability of symbolic execution of heap maninpulating software
%over GSE, while maintaining efficiency and precision.}
%\item{We prove that at any given point during symbolic execution, the symbolic 
%heap represents the exact set of feasible concrete heap structures 
%for the program under analysis.}
% \end{description}

%\nsr{needs
%  some text on the fact we use java lite to formalize the heap
%  summary, some key characteristicss of the heap such as the fact it
%  is a bipartiate graph with locations and referneces, properties of
%  the heap such as the determinism and immutability}

%\nsr{Is any part of the formalism borrowed
%from the dilligs? Which part are unique?}  \nsr{end on a couple
%sentences about the various rules that will be talked about in the
%paper}

%\nsr{The reader should be convinced about our expressiveness
%  claim here}

%\nsr{a paragraph that provides intution on how we are a bisumulation
%  with the GSE..reader should be convinced about the precision claim}

\begin{comment}
\newcommand{\Indstate}[1][1]{\State\hspace{#1\algorithmicindent}}

\begin{figure}
\begin{center}
\begin{algorithmic}[1]
\Procedure{load}{$r,f$}
\If{$r$ is initialized and $f$ is uninitialized}
\If{ $f$ has a reference type}
\State nondeterministically assign to $f$:
\State  \ \ null  %hacky way to indent.
\State \ \ non-null
\EndIf
\If{$f$ has a primitive type}
\State assign $f$ a new symbol
\EndIf
\EndIf
\If{$r$ is uninitialized}
\State nondeterministically assign to $f$:
\State  \ \ a new symbolic location  %hacky way to indent.
\State \ \ an existing symbolic location
\EndIf
\State\Return $r.f$
\EndProcedure
\end{algorithmic}
\end{center}
\caption{Lazier\# Initialization}
\end{figure}

Lazy initialization is a symbolic execution technique where memory structures are created on-demand, as they are required.
\end{comment}

>>>>

In recent years symbolic execution has provided the basis for various
software testing and analysis techniques. Symbolic execution explores
the program by treating input values as symbolic, and for each
explored path, it computes a set of feasible constraints on the
symbolic inputs to create a \emph{path condition}. Path conditions
computed by symbolic execution characterize the observed program
execution behaviors and have been used as an enabling technology for
various applications, e.g., regression
analysis~\cite{backes:2012,Godefroid:SAS11,Person:FSE08,person:pldi2011,Ramos:2011,Yang:ISSTA12},
data structure repair~\cite{KhurshidETAL05RepairingStructurally},
dynamic discovery of
invariants~\cite{CsallnerETAL08DySy,Zhang:ISSTA14}, and
debugging~\cite{Ma:2011}.

The path conditions computed by symbolic execution have two important
features: 1) they characterize the set of feasible execution paths,
i.e., infeasible paths are omitted, and 2) each path condition encodes
the set of concrete program inputs that would cause execution to
follow the same control flow path. Because path conditions are encoded
as constraints over program inputs, the reasoning capabilities of
symbolic execution are limited by the underlying constraint
solver. Recent work has explored the strengths and limitations of
tools which leverage constraint solvers, e.g., concolic execution,
identifying symbolic dereferencing and updating operations as areas
for potential improvement~\cite{Chen:2013,Qu:2011}.  Precisely
modeling these operations, however, is challenging because it requires
formulating logical predicates over an input domain that contains a
potentially unbounded number of references.

The most direct approach to address these challenges is to approximate
the results of a symbolic deference by substituting in a valid
concrete value, as is done in dynamic symbolic execution
(\dsetxt{})~\cite{Godefroid:POPL07,Godefroid:2005,Sen:2005,Tillmann:2008}. These
approaches partition the universe of heaps in the symbolic
representation but often under-approximate the partitions or require
array theories in the constraint
solver~\cite{Chen:2013,Elkarablieh:2009}. A more effective approach is
to partition the reachable heaps by using value summary
representations, where multiple values for a reference may be
represented along a single execution
path~\cite{Dillig:2011,Elkarablieh:2009,Sen:2014,Torlak:2014,Xie:2005,Yorsh:2008}. Systems
using these representations perform analyses that are more complete
than standard~\dsetxt{} methods, but their associated dereferencing
operations require placing further restrictions on the input heap,
such as limiting the number of nodes~\cite{Elkarablieh:2009}, the
nature of aliasing relationships~\cite{Babic:2007,Xie:2005}, or the
the use of recursive data structures~\cite{Dillig:2011}.

An alternative approach which does not under-approximate the heap is
Generalized Symbolic Execution
(\gsetxt{})~\cite{Cadar:2008,KiasanKunit,GSE03,Rosner:2015}.  On symbolic
dereferencing operations,~\gsetxt{} partitions the input space between the
various possible aliasing configurations on-the-fly and also creates
non-deterministic choice points on the current path in order to
explore each heap configuration. This approach, however, is
inefficient because it does not treat heap allocated data as symbolic,
but instead concretizes all possible heap configurations.

The contribution of this work is a non-trivial adaptation of concepts
in abstract interpretation to the domain of symbolic execution. The
result of which, unlike~\gsetxt{}, is the ability to efficiently generate
and update \emph{symbolic heaps} to produce a single path constraint
that is sound and complete up to the point of execution. This is
accomplished by representing the symbolic heap with a field sensitive
points-to graph with edges annotated with disjunctive formulas
describing aliasing relationships. Where~\gsetxt{} branches the heap at every
de-reference, the symbolic heap in this work splits only at reference
comparison operations or field operations that may throw exceptions.

The correctness of the symbolic heap and its on-the-fly update rules
for symbolic execution is shown by proving the existence of a
bisimulation between symbolic heap states and~\gsetxt{} states. From that
result comes the soundness and completeness of the path constraint up
to the point of execution.  The efficiency of the representation is
established with an exploratory empirical evaluation within the Java
PathFinder framework over a classical set of symbolic execution
benchmarks. The evaluation demonstrates how the symbolic heap scales
compared to state-of-art~\gsetxt{} algorithms.

This new symbolic heap and update rules for symbolic execution is a
novel approach for de-referencing and updating heap allocated data
that brings together the best features of previous work while removing
the limitations of that same work. Although it draws inspiration from
abstract interpretation by defining a symbolic heap,
\begin{compactitem}
  \item it is an exact representation, without over- or under-approximation;
  \item it imposes no restrictions on the set of heaps that can be defined;
  \item it does not require the creation of an initial set of input heaps;
  \item it is not tied to specific constraint solver theories, e.g., arrays; and
  \item it does not require global constraint rewriting for on-the-fly updates.
\end{compactitem}
These several properties set this work apart from other related works.

\begin{comment}
\section{Introduction}

% SymExe is cool because for reasons x,y, and z

In recent years symbolic execution has provided the basis for various
software testing and analysis techniques. Symbolic execution
systematically explores the program execution space representing input
values symbolically, and for each explored path, it computes
constraints on the symbolic inputs to create a \emph{ path condition}.
Path conditions computed by symbolic execution characterize the
observed program execution behaviors and have been used as an enabling
technology for various applications, e.g., regression
analysis~\cite{backes:2012,Godefroid:SAS11,Person:FSE08,person:pldi2011,Ramos:2011,Yang:ISSTA12},
data structure repair~\cite{KhurshidETAL05RepairingStructurally},
dynamic discovery of
invariants~\cite{CsallnerETAL08DySy,Zhang:ISSTA14}, and
debugging~\cite{Ma:2011}.

The path conditions computed by symbolic execution have two important
characteristics: 1) they characterize the set of feasible execution
paths, i.e., infeasible paths are omitted, and 2) each path condition
represents the set of concrete program inputs that would cause
execution to follow that program path. Because path conditions are
encoded as constraints over program inputs, the reasoning capabilities
of symbolic execution are limited by the underlying constraint solver.
Thus, extending the capabilities of symbolic execution to reason about
new theories is an area of active research. Of particular interest are
symbolic dereferencing and updating
operations~\cite{Chen:2013,Qu:2011}. Precisely modeling these
operations is challenging, because it requires formulating logical
predicates over an input domain containing a potentially unbounded
number of references.

Current strategies for symbolic dereferencing require making tradeoffs
between the completeness of the analysis and the completeness of the
path condition. For example, one simple approach is to terminate
execution upon attempting to dereference any symbolic
value~\cite{Clarke:2004}. While this preserves the integrity of the
path condition, it comes at the cost of an incomplete analysis. To
enable a more complete analysis, dynamic symbolic execution (\dsetxt{}),
approximates the results of a symbolic dereferencing by substituting
in a valid concrete
value~\cite{Godefroid:POPL07,Godefroid:2005,Sen:2005,Tillmann:2008}. This
allows execution to continue, but the approximation renders the path
condition incomplete~\cite{Chen:2013,Elkarablieh:2009}, by effectively
partitioning the space of input heaps into those where the result of
the dereferencing matches the concrete execution. Larger partitions
can be achieved by using value summary representations, where multiple
values for a reference may be represented along a single execution
path~\cite{Dillig:2011,Elkarablieh:2009,Sen:2014,Torlak:2014,Xie:2005,Yorsh:2008}. Systems
using these representations perform analyses that are more complete
than standard~\dsetxt{} methods, but their associated dereferencing
operations require placing further restrictions on the input heap,
such as limiting the number of nodes~\cite{Elkarablieh:2009}, the
nature of aliasing relationships~\cite{Babic:2007,Xie:2005}, or the
shape of the heap itself~\cite{Dillig:2011}. Finally, at the opposite
end of the spectrum lies generalized symbolic execution
(\gsetxt{})~\cite{Cadar:2008,KiasanKunit,GSE03,Rosner:2015}, which
accomplishes a fully complete analysis at the cost of an incomplete
path condition \sjp{not clear what an incomplete path condition
  is}. On symbolic dereferencing operations,~\gsetxt{} partitions the input
space between the various possible aliasing configurations
\sjp{so?}. Thus, none of the current techniques provide the option of
maintaining both a complete analysis and a complete path
condition. This is in stark contrast to symbolic execution over
primitive types, which enjoys both properties simultaneously.

\sjp{Need to explain that the techniques still operate within some bound for loops, etc.}

\sjp{Notion of on-the-fly updates is not here}

\sjp{Nowhere do we refer to the technique as a summary heap framework or algorithm,
so the next section with its heading just jumps out of nowhere.}

This paper introduces a technique that avoids these problems by
preserving the following properties of classical symbolic execution:
(i) Inputs are represented by unbounded and unconstrained input
symbols. This lack of predetermined constraints is key to avoiding
under-approximation, which in turn is required for performing a
complete analysis. (ii) Points of nondeterminism are only introduced
at control flow instructions, which is critical for maintaining a
complete path condition for a given control flow path.

@@

\item\textbf{-} The first sound and complete system for reasoning symbolically about the 
set of input heaps along any valid program path.

\item\textbf{-} A bisimulation proof establishing the soundness and 
completeness of the heap summary approach with respect to
properties provable by~\gsetxt{}.

\item\textbf{-} A proof-of-concept implementation and empirical study 
demonstrating the scalability of the summary heap approach
compared to other~\gsetxt{} approaches.

\end{compactdesc}
\end{comment}
